### 2018-09-27 paper.tex
Wip

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{hyperref}

@@

\begin{table}[h!]
  \begin{center}
    \caption{Traditional software vs reusable models.}
    \label{software-vs-models}
    \begin{tabular}{l|l}
      Traditional software & Installed code
      Reusable models & Installed code + trainable on novel data
    \end{tabular}
  \end{center}
\end{table}

The remainder of this paper will outline an approach to package
management applied to machine learning models that incorporates
\emph{model operations} and \emph{run management} as a central enabler
of model code reuse.

TODO:

- What problem is being solved?

Need to define reuse. Start with traditional softwar reuse and make it
clear that Guild addresses model reuse. What is model reuse?

- How package managers facilitate code reuse

Apply analogy of system packaging and the importance of formalizing
the process with protocols and reference implementations.

@@

- How machine learning models are distributed today (pip, GitHub
projects)

- Packaging tools

- Frameworks and libraries

- TensorFlow Hub

- Keras Applications

- ???

@@

- How is Guild different from these cited projects?

- Include model and operation definitions

- Reuse through distribution and patching

- Federation (protocol driven, intelligence moved to the perimeter).

@@

@@

@@

@@

\subsection{Operations}

@@

\subsection{Resources}

@@

@@

@@

\section{Use Cases}

\subsection{Publish a Model for Reuse}

@@

\subsection{Package Third Party Models}

@@

\subsection{Streamline Model Developer Workflow}

@@

\section{Managing Runs}

@@

>>>>

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{listings}

\lstset{
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  captionpos=b,
  framextopmargin=5pt,
  framexbottommargin=5pt,
  frame=tb
}

@@

The remainder of this paper will outline an approach to package that
incorporates \emph{model operations} and \emph{run management} as a
central enabler of code reuse for machine learning models.

@@

\subsubsection{Model zoos}

A common approach to sharing and reusing machine learning models is a
\emph{model zoo}---an online repository of reusable models that
includes source code, examples, documentation, and in some cases
pretrained models. The \href{Caffe Model Zoo}{Caffe Model Zoo} was one
of the first model zoos and has inspired serveral others (ref).

Model zoos encourage code reuse by consolidating related models
(e.g. based on underlying library implementation or application
domain) and providing consistent documentation and repeatable methods
for using the models.

Model zoos, however, do not typically provide tools for packaging,
distributing, and using models, beyond those of publishing source code
and documentation (e.g. GitHub, web sites, etc.)

\subsubsection{TensorFlow Hub and Keras Applications}

\href{https://www.tensorflow.org/hub/}{TensorFlow Hub modules} and
\href{https://keras.io/applications/}{Keras Applications} are examples
of framework level facilities that provide a higher degree of model
reuse vis-\'a-vis model zoos. Both TensorFlow Hub modules and Keras
application support tools and patterns for packaging, distributing,
and using model source code and associated pretrained models. They
emphasize the use of \emph{transfer learning} where applicable for
training models on novel datasets.

\subsubsection{Software packages and source coderepositories}

Models are commonly reused through traditional software distribution
channels: packages and source code repositories. In both cases, model
source code is provided along with varying levels of documentation and
supporting tools. Unlike model zoos and framework packages, there is
little consistency in quality of code reuse available through these
channels. Developers are left to their own accord to find and use
available models.

@@

Guild AI combines elements from model zoos, framework packages, and
traditional software channels to provide a state of the art facility
for packaging, distributing and using machine learning models.

As with model zoos, users can discover Guild AI models, including
documentation on their use. Guild AI additionally supports
installation and direct use of models through operations.

Guild AI superficially resembles TensorFlow Hub in that users can
discover, install, and use models quickly and easily. Guild AI however
differs considerably in its implementation. Guild AI does not present
a language level interface but instead defines a set of conventions
that are supported through configuration. Guild AI packaging
facilities can be applied to any machine learning model
implementation.

Guild AI also supports the well established practice of package
maintenence, where a third party can create usable packages without
necessarily coordinating with model developers. This is discussed in
detail in Section \ref{sec:package-maintainer}.

As with traditional software packages and source code repositories,
Guild AI supports a federated approach to software distriution. Users
are free to use Guild AI to create and distribute reusable models in a
variety of ways. Refer to Section \ref{sec:use-cases} for coverage of
various use case.

@@

This paper relies on a number of core concepts, which are defined
below.

@@

A \emph{project} is a directory that contains model related
code. Model developers maintain projects for their models. Model users
maintain projects for the code that works with models, possibly
training them or otherwise incorporating them into an application.

Projects are central to model development and use. Guild AI supports
project based workflow as described in Section
\ref{sec:streamline-workflow}.

\subsection{Guild File}

A \emph{Guild file} is a plain text file containing Guild AI related
configuration. The file is named \verb|guild.yml| and is typically
located in the root directory of a Guild AI enabled project.

Figure \ref{fig:guild-file} illustrates a simple Guild file.

\begin{figure}
{\footnotesize
\begin{verbatim}
 model: mnist
 description:
   Logistic regression classifier
   of MNIST digits
 operations:
   train:
     main: train_mnist
     flags:
       batchs-size: 32
       train-steps: 1000
   evaluate:
     main: eval_mnist
\end{verbatim}}
\caption{Guild file example}
\label{fig:guild-file}
\end{figure}

@@

A Guild AI \emph{package} is a Python wheel distribution (ref) that
was generated using the \verb|guild package| command. Guild AI
packages are no different from standard Python wheel distributions and
are generated using \verb|setuptools|. However, they contain Guild AI
specific entry points that effectively publish the models they
contain.

@@

A \emph{model} in Guild AI is a representation of a machine learning
model that is defined in a Guild file. Models define \emph{operations}
(Section \ref{sec:operations}) that perform model related tasks such
as training or evaluating as well as \emph{resources} (Section
\ref{sec:resources}) that operations may require.

Models may be defined according to the underlying architecture or
intended use of the model. Guild AI does not prescribe any formal
meaning to the term \emph{model} or what operations are supported.

\subsection{Operations}
\label{sec:operations}

An \emph{operation} is an action that can be performed on a
model. Operations are part of a model definition.

Guild AI does not prescribe any particular set of operations that must
be defined for a model. Model developers are free to name operations
as they see fit, though Guild AI does encourage naming conventions.

The \verb|train| operation---the act of training a model from
scratch---for example, is an operation that most models
support. However, the \verb|transfer-learn| operation---the act of
transfering knowledge learned in one domain to another domain---is a
form of training, but frequently appears in Guild AI models as an
operation distinct from \verb|train|.

@@

\emph{Flags} are parameters that can be set or modified by users when
they run an operation. Flags are used as inputs to the underlying code
that implements an operation.

\subsection{Resources}
\label{sec:resources}

A Guild AI \emph{resource} represents a set of source files that can
be made available as inputs to a model operation. Resources play a key
role in Guild AI's ability to effective use models. Resources can be
used to resolve project files, download files from URLs, unpack
archives, and reference files generated by other operation.

Figure \ref{fig:resource} illustrates a resource definition for a
model. The resource \verb|idx-data| defines four source files that
will be made available to any operation that requires that resource.

\begin{figure}
{\footnotesize
\begin{verbatim}
 model: mnist
 resources:
   idx-data:
     description:
       MNIST dataset in compressed
       IDX format
     sources:
       - train-images-idx3-ubyte.gz
       - train-labels-idx1-ubyte.gz
       - t10k-images-idx3-ubyte.gz
       - t10k-labels-idx1-ubyte.gz
\end{verbatim}}
\caption{Model resource example}
\label{fig:resource}
\end{figure}

@@

A \emph{run} is a file system artifact---specifically a directory
created and managed by Guild AI---that contains metadata and files
generated by an operation. A run may also refer to the operation
system process associated with a running operation.

Runs are described in detail in Section \ref{sec:runs}.

@@

Guild AI supports \emph{environments}, which are isolated runtime
environments used for installing and using models. Guild AI supports
environments created using
\href{https://virtualenv.pypa.io}{\emph{virtualenv}}.

@@

Guild AI \emph{workflows} are patterns of model use that are enabled
by running operations in specific order. Worflows are not currently
formal constructs in Guild AI but are referenced by convention or
documented usage.

\section{Use Cases}
\label{sec:use-cases}

The package management facilities in Guild AI enable a number of use
cases, which are described below. In each case, the following roles
may be referenced:

\begin{description}
    \item[user] person using a model developed by someone else
    \item[developer] person developing a model, either from scratch or
      adapting the work of others
    \item[packager] person neither using nor developing a model but
      who packages models from developers for use by users
\end{description}

\subsection{Publish a Model for Use}

A developer may encourage others to use her model by documenting the
steps needed to install, adapt, and use it. She may be motivated to
simplify the user experience so that her work is more easily and
succesfully applied. The benefit to her may be quality feedback from
expert users, improved reputation, a sense of contribution, or her
colleagues may simply expect that work be documented and easy to use.

While documenting the steps needed to use a model is undoubtedly
helpful, the process is still burdensome for both the developer and
her users. Guild AI can be used to simultaenously document model use
but also directly support its use through model operations.

Consider these instructions for model use, which might be documented
for a project:

{\footnotesize
\begin{verbatim}
1. Install required packages
2. Download the raw data
3. Prepare the raw data for training
4. Train the model
5. Export the trained model for deployment
\end{verbatim}}

While these steps provide invaulable information, they still rely on
the user to carefully execute them and troubleshoot issues as they
arise. The process, while thoroughly documented, can take time and is
subject to error at any step.

Guild AI can be used to both document and automate these
steps. Listing~\ref{lst:published-model} illustrates a Guild file that
might be used to codify the steps as model operations, which are both
discoverable and runnable.

\begin{lstlisting}[caption=whatever,label={lst:published-model}]
model: sample
operations:
  prepare-data:
    description:
      Prepare data for training
    main: prepare_data
    requires: raw-data
  train:
    description: Train model
    main: train
    requires: prepared-data
  export:
    description:
      Export model for deployment
    main: export
    requires: trained-model
resources:
  raw-data:
    description: Raw data for training
    sources:
      - url: http://mydata.com/raw.tgz
  prepare-data:
    description: Prepared data
    sources:
      - operation: prepare-data
  trained-model:
    description: Trained model
    sources:
      - operation: train
\end{lstlisting}

@@

\subsection{Package Third Party Models}
\label{sec:package-maintainer}

@@

\subsection{Streamline Model Developer Workflow}
\label{sec:streamline-workflow}

@@

\section{Managing Runs}
\label{sec:runs}

@@

\begin{table}[h!]
  \begin{center}
    \caption{Traditional software vs reusable models.}
    \label{software-vs-models}
    \begin{small}
      \begin{tabular}{l|l}
        \toprule
        \textbf{Traditional software} & Installed code \\ \hline
        \textbf{Reusable models} & Installed code + trainable on novel data \\
        \bottomrule
      \end{tabular}
    \end{small}
  \end{center}
\end{table}
